---
title: "Covid91 vaccine study Final2024"
author: "Gianna Chevez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Is the Covid vaccine effective  in males?

$H_0$: The new preventative vaccine (Covid921) does not exhibit any effectiveness  on males.  
$H_A$: The new preventative vaccine (Covid921) does exhibit effectiveness on male.

cat-cat

```{r}
Males = FinalData[FinalData$sex=='M',]
head(Males)
```


## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```


The graph indicates that the vaccine is effective in preventing COVID-19 among males, as a lower proportion of vaccinated men contracted the virus compared to those who received the placebo.


## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```


The table shows that 26.20% of men who received the vaccine contracted COVID-19, whereas 73.80% of men who received the placebo became infected.


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The p-value from the test is 0.00000000000000022, which is less than 0.05, so we must reject the null hypothesis. The odds ratio from Fisher’s exact test was 2.815, indicating that the vaccine is approximately 2.8 times more effective in men.

```{r}
guysfish <- fisher.test(table1)
guysfish
```


```{r}
low <- (guysfish$conf.int[1] / (guysfish$conf.int[1] + 1)) * 100
hi <- (guysfish$conf.int[2] / (guysfish$conf.int[2] + 1)) * 100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The p-value from Fisher’s exact test is effectively zero—well below the conventional 0.05 threshold. As a result, we reject the null hypothesis, providing strong evidence that the vaccine is effective in males.

The odds ratio is 2.81, indicating that men who received the vaccine were nearly three times less likely to contract COVID-19 compared to those who received the placebo. Additionally, the 95% confidence interval ranges from 2.21 to 3.61. Since the entire interval is above 1, this further supports the conclusion that the vaccine significantly reduces the risk of infection.

Overall, the data provide compelling evidence that the vaccine is a beneficial and effective intervention for men.


## Females

##Methods  

$H_0$: The preventative vaccine does not exhibit its effective in females        
$H_A$: The preventative vaccine exhibits its effective in females.  
Cat~Cat.

```{r}
Females = FinalData[FinalData$sex=='F',]
head(Females)
```

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

The bar graph indicates that the vaccine is more effective in women, as a smaller proportion of vaccinated women contracted COVID-19 compared to those who received the placebo.


##Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that approximately 35.07% of women who received the vaccine contracted COVID-19, compared to 64.93% of women who received the placebo. This indicates that the vaccine is effective in reducing COVID-19 cases among women.


##Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The p-value from the test is 0.00000001914, which is well below the 0.05 significance threshold. Therefore, we reject the null hypothesis, providing strong evidence of the vaccine’s effectiveness in women. The odds ratio from Fisher’s exact test is 1.856, indicating that women who received the vaccine were approximately 1.86 times less likely to contract COVID-19 compared to those who received the placebo.



```{r}
guysfish <- fisher.test(table1)
guysfish
```

```{r}
low <- (guysfish$conf.int[1] / (guysfish$conf.int[1] + 1)) * 100
hi <- (guysfish$conf.int[2] / (guysfish$conf.int[2] + 1)) * 100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The p-value from Fisher’s exact test was 1e-08, which is well below the standard 0.05 threshold. As a result, we reject the null hypothesis, indicating that the vaccine is effective for females as well. The odds ratio is 1.9, suggesting that vaccinated women were nearly twice as likely to avoid contracting COVID-19 compared to those who received the placebo.

Additionally, the 95% confidence interval ranges from 1.5 to 2.3. Because the entire interval lies above 1, it provides further evidence that the vaccine has a statistically significant protective effect. Overall, these findings suggest that the vaccine is a beneficial and effective option for women.


## LGBTQ

##Methods  
$H_0$ The preventative vaccine does not exhibit its effective in people who are LGBTQ.        
$H_A$ The preventative vaccine exhibits its effective in people who are LGBTQ. 

Cat~Cat

```{r}
gay = FinalData[FinalData$LGBTQ=='gay',]
head(gay)
```


#Graphical Results 

```{r}
barchartGC(~infected + treatment,data=gay)
barchartGC(~infected + treatment,data=gay, type="percent")
```

The bar graph indicates that the vaccine is effective in preventing COVID-19 among individuals who identify as LGBTQ.


##Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The table shows that 20% of LGBTQ individuals who received the vaccine contracted COVID-19, compared to 80% of those who received the placebo. This suggests that the vaccine is effective in reducing COVID-19 infections within the LGBTQ community.


##Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The p-value from the test is 0.000002266, which is well below the 0.05 significance threshold. Therefore, we reject the null hypothesis, indicating strong evidence that the vaccine is effective. The odds ratio from Fisher’s exact test is 4.27, suggesting that individuals who received the vaccine were approximately 4.27 times more likely to avoid contracting COVID-19 compared to those who received the placebo.



## Druggies

##Methods  

$H_0$ The preventative vaccine does not exhibit its effective in people who are drug users.    
$H_A$ The preventative vaccine exhibits its effective in people who are drug users.  

Cat~Cat

```{r}
DrugUser = FinalData[FinalData$DrugUser=='yes',]
```


##Graphical Results

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```

The bar graph shows that the vaccine appears to be less effective in individuals who use drugs.Add comment



##Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```



The table shows that 91.67% of drug users who contracted COVID-19 had also received the vaccine. While this may initially suggest that the vaccine is less effective among individuals who use drugs, this observation alone does not account for factors such as sample size, overall vaccination rates in the group, or potential confounding variables. Further statistical analysis would be needed to draw reliable conclusions about vaccine effectiveness in this subgroup.



##Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The p-value from the chi-squared test is 0.0000000001414, which is far below the 0.05 threshold, so we reject the null hypothesis. However, despite this strong statistical significance, the results indicate that the vaccine may have limited effectiveness in this particular group.


```{r}
guysfish <- fisher.test(table1)
guysfish
```

```{r}
low <- (guysfish$conf.int[1] / (guysfish$conf.int[1] + 1)) * 100
hi <- (guysfish$conf.int[2] / (guysfish$conf.int[2] + 1)) * 100
options(digits=2)
```



# Overall Results and Conclusions

Effectiveness of the Vaccine Across Different Groups
Men:
Based on the data, the vaccine appears to be highly effective for men. The table shows that only 26.2% of men who received the vaccine contracted COVID-19, compared to 73.8% of men who received the placebo—a substantial and clinically meaningful difference. The p-value associated with this comparison is 2.2 × 10⁻¹⁶, which is far below the 0.05 significance threshold, leading us to reject the null hypothesis. This provides very strong evidence that the vaccine has a real protective effect in men. The odds ratio of approximately 2.8 indicates that vaccinated men were about 2.8 times more likely to avoid infection compared to men who received the placebo. Overall, these results suggest the vaccine is a reliable and effective preventive measure for men.


Women: 
The vaccine also shows clear effectiveness for women. Among women who contracted COVID-19, 35.07% had received the vaccine, whereas 64.93% had received the placebo, reflecting a meaningful reduction in infection rates due to vaccination. The p-value for this group was 1.914 × 10⁻⁸, which is well below 0.05, allowing rejection of the null hypothesis and supporting the vaccine’s benefit. The odds ratio of approximately 1.86 suggests that vaccinated women were significantly less likely to contract COVID-19 than unvaccinated women. Thus, the vaccine appears to be an effective preventive option for women as well.

LGBTQ Individuals:
For LGBTQ individuals, the data indicate strong vaccine effectiveness. Only 20% of LGBTQ people who contracted COVID-19 had been vaccinated, compared to 80% who received the placebo, demonstrating a substantial protective effect. The p-value of 2.266 × 10⁻⁶ is far below the 0.05 threshold, again rejecting the null hypothesis. The odds ratio from Fisher’s exact test was 4.27, meaning vaccinated LGBTQ individuals were more than four times as likely to avoid COVID-19 infection compared to those who received the placebo. These findings suggest the vaccine is highly effective for this group.


Individuals Who Use Drugs:
In contrast, the vaccine appears to be less effective among individuals who use drugs. The table shows an unexpected pattern: 91.67% of drug users who contracted COVID-19 had been vaccinated, while only 8.33% had received the placebo. This inverse trend suggests limited vaccine protection in this subgroup. The p-value was 1.414 × 10⁻¹⁰, which is statistically significant, so we reject the null hypothesis; however, the outcome is opposite to that seen in other groups. The odds ratio of 0.087 indicates that vaccinated drug users were much less likely to avoid infection, implying that the vaccine may not provide strong protection for individuals who use drugs. Further research is needed to understand the underlying reasons for this finding, such as potential confounders, differences in immune response, or behavioral factors.

Summary:
Across most groups—men, women, and LGBTQ individuals—the vaccine demonstrates strong, statistically significant effectiveness in reducing COVID-19 infections. However, the data suggest that vaccine effectiveness may be limited among drug users, highlighting the need for targeted investigation and possibly additional interventions for this population.



