---
title: "Covid91 vaccine study Final2024"
author: "Gianna Chevez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Is the Covid vaccine effective  in males?

$H_0$: The new preventative vaccine (Covid921) does not exhibit any effectiveness  on males.  
$H_A$: The new preventative vaccine (Covid921) does exhibit effectiveness on male.

cat-cat

```{r}
Males = FinalData[FinalData$sex=='M',]
head(Males)
```


## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```


The graph indicates that the vaccine is effective in preventing COVID-19 among males, as a lower proportion of vaccinated men contracted the virus compared to those who received the placebo.




## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions