---
title: "Covid91 vaccine study Final2024"
author: "Gianna Chevez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Is the Covid vaccine effective  in males?

$H_0$: The new preventative vaccine (Covid921) does not exhibit any effectiveness  on males.  
$H_A$: The new preventative vaccine (Covid921) does exhibit effectiveness on male.

cat-cat

```{r}
Males = FinalData[FinalData$sex=='M',]
head(Males)
```


## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```


The graph indicates that the vaccine is effective in preventing COVID-19 among males, as a lower proportion of vaccinated men contracted the virus compared to those who received the placebo.


## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```


The table shows that 26.20% of men who received the vaccine contracted COVID-19, whereas 73.80% of men who received the placebo became infected.


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The p-value from the test is 0.00000000000000022, which is less than 0.05, so we must reject the null hypothesis. The odds ratio from Fisher’s exact test was 2.815, indicating that the vaccine is approximately 2.8 times more effective in men.

```{r}
guysfish <- fisher.test(table1)
guysfish
```


```{r}
low <- (guysfish$conf.int[1] / (guysfish$conf.int[1] + 1)) * 100
hi <- (guysfish$conf.int[2] / (guysfish$conf.int[2] + 1)) * 100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The p-value from Fisher’s exact test is effectively zero—well below the conventional 0.05 threshold. As a result, we reject the null hypothesis, providing strong evidence that the vaccine is effective in males.

The odds ratio is 2.81, indicating that men who received the vaccine were nearly three times less likely to contract COVID-19 compared to those who received the placebo. Additionally, the 95% confidence interval ranges from 2.21 to 3.61. Since the entire interval is above 1, this further supports the conclusion that the vaccine significantly reduces the risk of infection.

Overall, the data provide compelling evidence that the vaccine is a beneficial and effective intervention for men.


## Females

##Methods  

$H_0$: The preventative vaccine does not exhibit its effective in females        
$H_A$: The preventative vaccine exhibits its effective in females.  
Cat~Cat.

```{r}
Females = FinalData[FinalData$sex=='F',]
head(Females)
```

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

The bar graph indicates that the vaccine is more effective in women, as a smaller proportion of vaccinated women contracted COVID-19 compared to those who received the placebo.


##Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The numerical results show that approximately 35.07% of women who received the vaccine contracted COVID-19, compared to 64.93% of women who received the placebo. This indicates that the vaccine is effective in reducing COVID-19 cases among women.


##Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The p-value from the test is 0.00000001914, which is well below the 0.05 significance threshold. Therefore, we reject the null hypothesis, providing strong evidence of the vaccine’s effectiveness in women. The odds ratio from Fisher’s exact test is 1.856, indicating that women who received the vaccine were approximately 1.86 times less likely to contract COVID-19 compared to those who received the placebo.



```{r}
guysfish <- fisher.test(table1)
guysfish
```

```{r}
low <- (guysfish$conf.int[1] / (guysfish$conf.int[1] + 1)) * 100
hi <- (guysfish$conf.int[2] / (guysfish$conf.int[2] + 1)) * 100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The p-value from Fisher’s exact test was 1e-08, which is well below the standard 0.05 threshold. As a result, we reject the null hypothesis, indicating that the vaccine is effective for females as well. The odds ratio is 1.9, suggesting that vaccinated women were nearly twice as likely to avoid contracting COVID-19 compared to those who received the placebo.

Additionally, the 95% confidence interval ranges from 1.5 to 2.3. Because the entire interval lies above 1, it provides further evidence that the vaccine has a statistically significant protective effect. Overall, these findings suggest that the vaccine is a beneficial and effective option for women.


## LGBTQ

##Methods  
$H_0$ The preventative vaccine does not exhibit its effective in people who are LGBTQ.        
$H_A$ The preventative vaccine exhibits its effective in people who are LGBTQ. 

Cat~Cat

```{r}
gay = FinalData[FinalData$LGBTQ=='gay',]
head(gay)
```


#Graphical Results 

```{r}
barchartGC(~infected + treatment,data=gay)
barchartGC(~infected + treatment,data=gay, type="percent")
```

The bar graph indicates that the vaccine is effective in preventing COVID-19 among individuals who identify as LGBTQ.


##Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The table shows that 20% of LGBTQ individuals who received the vaccine contracted COVID-19, compared to 80% of those who received the placebo. This suggests that the vaccine is effective in reducing COVID-19 infections within the LGBTQ community.




## Druggies


# Overall Results and Conclusions